### 2015-10-29 manuscript.Rnw
some spelling

The R programming language is a popular tool for  phylogenetics and comparative
analysis. The R packages \texttt{ape} \citep{Paradis2004}, \texttt{phylobase}
\citep{phylobase-0.8.0}, and \texttt{RNeXML} \citep{Boettiger2015} each provide
functions to import and manipulate trees within R and save the results in
standard data formats. Additional packages including \texttt{phytools}
\citep{Revell2012}, \texttt{geiger} \citep{Pennell2014-geiger}, and \texttt{ggtree} 
\citep{Yu-ggtree} allow users to analyse and visualise data in a 
phylogenetic context (see 
\url{https://cran.r-project.org/web/views/Phylogenetics.html} for a
comprehensive list of phylogenetics packages in R). In addition to packages for
phylogentic and comparative analyses, a growing number of R packages allow users to query and
access data from the web (e.g. \texttt{rFISHBASE} \citep{Boettiger2012}, \texttt{rAvis} \citep{Varela2014}, and \texttt{paleoDB} \citep{Varela2015}), meaning data 
associated with taxa in a given phylogeny can be obtained directly in R. 
These packages, combined with the language's support for literate programming 
(\citealt{Knuth1984}; \citealt{Xie2014}), make R an ideal platform for reproducible research
in phylogenetics and comparative biology, as they allow a complete record of the 
steps taken in gathering, processing, and analysing a given data set to be
produced.

Here we present \texttt{rotl}, and R package that allows users to download
phylogenetic and taxonomic data from the OTL directly in R. \texttt{rotl} takes
advantage of Application Programming Interfaces (APIs) provided by OTL to access
sub-trees from the synthetic Open Tree, as well as the published source trees that 
contribute to the synthesis. By providing direct access to high quality
phylogenetic data in R, \texttt{rotl} fills a key gap typical comparative
analysis workflows, and extends the degree to which R supports reproducable
research in phylogenetics and comparative biology.

@@

The package is well-documented, and includes two package
vignettes (documents that demonstrate the use of the package and contain executable 
R code)
%as well as per-function documentation
. There is also an extensive
test-suite that covers both the internal functions that \texttt{rotl} uses to
connect to OTL and public functions users apply to acess and process data.

>>>>

The R programming language is a popular tool for  phylogenetics and comparative
analysis. The R packages \texttt{ape} \citep{Paradis2004}, \texttt{phylobase}
\citep{phylobase-0.8.0}, and \texttt{RNeXML} \citep{Boettiger2015} each provide
functions to import and manipulate trees within R and save the results in
standard data formats. Additional packages including \texttt{phytools}
\citep{Revell2012}, \texttt{geiger} \citep{Pennell2014-geiger}, and \texttt{ggtree} 
\citep{Yu-ggtree} allow users to analyze and visualize data in a 
phylogenetic context (see 
\url{https://cran.r-project.org/web/views/Phylogenetics.html} for a
comprehensive list of phylogenetics packages in R). In addition to packages for
phylogenetic and comparative analyses, a growing number of R packages allow users to query and
access data from the web (e.g. \texttt{rFISHBASE} \citep{Boettiger2012}, \texttt{rAvis} \citep{Varela2014}, and \texttt{paleoDB} \citep{Varela2015}), meaning data 
associated with taxa in a given phylogeny can be obtained directly in R. 
These packages, combined with the language's support for literate programming 
(\citealt{Knuth1984}; \citealt{Xie2014}), make R an ideal platform for reproducible research
in phylogenetics and comparative biology, as they allow a complete record of the 
steps taken in gathering, processing, and analyzing a given data set to be
produced.

Here we present \texttt{rotl}, and R package that allows users to download
phylogenetic and taxonomic data from the OTL directly in R. \texttt{rotl} takes
advantage of Application Programming Interfaces (APIs) provided by OTL to access
sub-trees from the synthetic Open Tree, as well as the published source trees that 
contribute to the synthesis. By providing direct access to high quality
phylogenetic data in R, \texttt{rotl} fills a key gap typical comparative
analysis workflows, and extends the degree to which R supports reproducible
research in phylogenetics and comparative biology.

@@

The package is well-documented, and includes two package
vignettes (documents that demonstrate the use of the package and contain executable 
R code)
%as well as per-function documentation
. There is also an extensive
test-suite that covers both the internal functions that \texttt{rotl} uses to
connect to OTL and public functions users apply to access and process data.
